Preparación de datos

### 2025-01-09 production/prod_prep.Rmd
change income vars

```{r}
remove(list = ls()) # clean workspace
# .rs.restartR() #restart R session
if (!require("pacman")) install.packages("pacman") # install pacman
# load libraries
pacman::p_load(dplyr,readxl,sjmisc, sjlabelled, questionr, car,here)
load(here::here("input/data-original/elsoc_long.Rdata"))

@@

# Attitudes toward inmigrants__________________________________________________
elsoc_long_2016_2023 <- 
elsoc_long_2016_2023 %>% 
  #create new variables
  mutate(just_educ=d02_02,
         just_pension=d02_01,
         just_salud=d02_03,
         des_perc=c18_11,
         sexo=m0_sexo,
         merit_effort=c18_09,
         merit_talent=c18_10,
         ahead_family=r12_02,
         ahead_educ=r12_03,
         ahead_ambition=r12_04,
         ahead_hardwork=r12_05,
         covid_conocidos=r03_09,
         covid_economia=c37_09,
         covid_cuarentena=c37_10,
         covid_propio=s31,
         covid_alimentos=m62,
         covid_beneficios=m63_01,
         covid_retirouno=m63_02,
         covid_retirodos=m63_03
         ) %>% 
  sjlabelled::drop_labels(., drop.na = FALSE)

#
sjmisc::frq(elsoc_long_2016_2023$just_educ)
elsoc_long_2016_2023$just_educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$just_educ, 
                      label = "Education distributive justice")

@@

sjmisc::frq(elsoc_long_2016_2023$just_educ)
sjmisc::frq(elsoc_long_2016_2023$merit_effort)
sjmisc::frq(elsoc_long_2016_2023$merit_talent)
sjmisc::frq(elsoc_long_2016_2023$ahead_family)
sjmisc::frq(elsoc_long_2016_2023$ahead_educ)
sjmisc::frq(elsoc_long_2016_2023$ahead_ambition)
sjmisc::frq(elsoc_long_2016_2023$ahead_hardwork)
# Socioeconomic_________________________________________________________________

# Education_______________________________________
elsoc_long_2016_2023$educ <- 
  car::recode(elsoc_long_2016_2023$m01,
              "c(1,2,3)=1;c(4,5)=2;c(6,7)=3;c(8,9,10)=4; c(-888,-999)=NA")
elsoc_long_2016_2023$educ <-
  factor(elsoc_long_2016_2023$educ,
         labels = c("Primary","High school","Technical","Universitary"))

class(elsoc_long_2016_2023$educyear)
sjmisc::frq(elsoc_long_2016_2023$educyear)

#Impute midpoint of income ranges
elsoc_long_2016_2023$m30_rec <-
as.numeric(car::recode(elsoc_long_2016_2023$m30,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

#Impute midpoint of income ranges (2021)
elsoc_long_2016_2023$m30b_rec <-
as.numeric(car::recode(elsoc_long_2016_2023$m30b,
           "1=125000;2=300000;3=400000;4=575000;5=70000;NA=NA;c(-888,-999)=NA"))

sjmisc::frq(elsoc_long_2016_2023$m30b_rec)

#Recode DK/DA of Income to NA
elsoc_long_2016_2023$m29_rec <-
  as.numeric(car::recode(elsoc_long_2016_2023$m29,"c(-888,-999)=NA"))

#replace NA of income with new imputed variable
elsoc_long_2016_2023$m29_imp <- 
  ifelse(test = !is.na(elsoc_long_2016_2023$m29_rec),
         yes =  elsoc_long_2016_2023$m29_rec,
         no =  elsoc_long_2016_2023$m30_rec)
summary(elsoc_long_2016_2023$m29_imp)

elsoc_long_2016_2023$m29_imp <- 
  ifelse(test = is.na(elsoc_long_2016_2023$m29_imp),
         yes =  elsoc_long_2016_2023$m30b_rec,
         no =  elsoc_long_2016_2023$m29_imp)
summary(elsoc_long_2016_2023$m29_imp)

# deflate at each year's prices
elsoc_long_2016_2023$deflactor <-
  with(elsoc_long_2016_2023, case_when(
    ola == 2016 ~ 113.88 / 123.82,
    ola == 2017 ~ 116.46 / 123.82,
    ola == 2018 ~ 119.45 / 123.82,
    ola == 2019 ~ 123.82 / 123.82
  ))

# N Household:
elsoc_long_2016_2023 <-
  elsoc_long_2016_2023 %>%
  mutate(n_hogar =
           dplyr::case_when(ola == 1 ~ nhogar1,
                            ola == 2 ~ m46_nhogar,
                            ola == 3 ~ m54,
                            ola == 4 ~ m54,
                            ola == 5 ~ m54,
                            ola == 6 ~ m54,
                            ola == 7 ~ m54))
sjmisc::frq(elsoc_long_2016_2023$n_hogar)

#Recode DK/DA to NA
elsoc_long_2016_2023$n_hogar_r<-
  car::recode(elsoc_long_2016_2023$n_hogar,"c(-888,-999)=NA")

# Per capita household income:
elsoc_long_2016_2023$ing_pc <- 
  (elsoc_long_2016_2023$m29_imp/elsoc_long_2016_2023$n_hogar_r)

elsoc_long_2016_2023$ing_pc <-
sjlabelled::set_label(x = elsoc_long_2016_2023$ing_pc,
                      label = "Household income per capita")

sjmisc::descr(elsoc_long_2016_2023$ing_pc)

# Compute income quintiles
elsoc_long_2016_2023 <- elsoc_long_2016_2023 %>% 
  group_by(ola) %>% 
  mutate(quintil = ntile(-desc(ing_pc), 5)) %>% 
  ungroup()

elsoc_long_2016_2023$quintil <- 
  factor(elsoc_long_2016_2023$quintil,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

elsoc_long_2016_2023$quintil <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$quintil,
                      label = "Household income quintile per capita")

sjmisc::frq(elsoc_long_2016_2023$quintil)

#include new quintile category with missing cases
elsoc_long_2016_2023$quintil1<-
  car::recode(elsoc_long_2016_2023$quintil, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

elsoc_long_2016_2023$quintil1 <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$quintil1,
                      label = "Household income quintile per capita (NA)") 
sjmisc::frq(elsoc_long_2016_2023$quintil1)

#Age Ranges
elsoc_long_2016_2023$edad <- 
  factor(car::recode(elsoc_long_2016_2023$m0_edad, 
                     "18:29=1;30:49=2;50:64=3;65:150=4"),
         labels = c('18-29', '30-49', '50-64', '65 or more'))
elsoc_long_2016_2023$edad <-
  sjlabelled::set_label(elsoc_long_2016_2023$edad, 
                        label = c("Age groups")) 
# Other controls______________________________________________________________
# Subjective social status: individual ____________________________

@@

```{r}
# Select variables______________________________________________________________

df_study1 <- 
elsoc_long_2016_2023 %>% 
  filter(muestra == 1) %>% #keep original sample (vs refresh)
  select(idencuesta,ola,region,region_cod,comuna,comunacod=comuna_cod,
         muestra,
         ponderador_long_total, segmento, estrato,
         sexo,
         # cuestion_mig,
         # starts_with("in_"),
         just_educ,
         just_pension,
         just_salud,
         des_perc,
         merit_effort,
         merit_talent,
         ahead_family,
         ahead_educ,
         ahead_ambition,
         ahead_hardwork,
         educ,educyear,
         # ing_pc,
         # quintil,
         quintil1,
         ess,
         covid_conocidos,
         covid_economia,
         covid_cuarentena,
         covid_propio,
         covid_alimentos,
         covid_beneficios,
         covid_retirouno,
         covid_retirodos,
         # ess_inm,
         # nation_t1,
         # know_inm,
         # know_inm_bi,
         # frien_inm_bi_t4,
         # frien_inm_bi_t1,
         # know_inm_t1t4,
         # frien_inm,
         # frien_inm_bi,
         # frien_inm_t1t4,
         pos_id,
         edad)

# Reshape long to wide
df_study1_wide <- df_study1 %>% 
  tidyr::pivot_wider(id_cols = c("idencuesta","muestra"),
                     names_from = "ola",
                     # names_prefix = c("educ","educyear","ing_pc","quintil","quintil1"),
                     values_from = names(select(df_study1,comuna,comunacod,ponderador_long_total,segmento, estrato, just_educ:edad,sexo))
                     )

@@

sjPlot::view_df(df_study1_wide,
                show.frq = T,show.values = T,show.na = T,show.prc = T, show.type = T)

# reshape from long to wide
pacman::p_load(datasets,data.table)
df_study1_long <- data.table::melt.data.table(data.table::setDT(df_study1_wide),
              id.vars = c("idencuesta","muestra"),
              variable.name = c("ola"),
              measure = patterns("^comuna_","^comunacod_","^ponderador_long_total", "^segmento", "^estrato",
                                 "^just_educ_", "^just_pension_", "^just_salud_", "^des_perc_",
                                 "^merit_effort_","^merit_talent_",
                                 "^ahead_family_", "^ahead_educ_", "^ahead_ambition_", "^ahead_hardwork_",
                                 "^covid_conocidos_", "^covid_economia_", "^covid_cuarentena_", "^covid_propio_",
                                 "^covid_alimentos_", "^covid_beneficios_", "^covid_retirouno_", "^covid_retirodos_",
                                 "^educ_","^educyear_","^quintil1_","^ess_",
                                 "^pos_id_",
                                 "^edad_","^sexo_"
                                 ),
              value.name = c("comuna","comunacod","ponderador_long_total","segmento","estrato",
                             "just_educ", "just_pension","just_salud", "des_perc",
                             "merit_effort","merit_talent",
                             "ahead_family","ahead_educ","ahead_ambition","ahead_hardwork",
                             "covid_conocidos", "covid_economia", "covid_cuarentena", "covid_propio",
                             "covid_alimentos", "covid_beneficios", "covid_retirouno", "covid_retirodos",
                             "educ","educyear","quintil1","ess",
                             "pos_id",
                             "edad","sexo"),
              na.rm = F,value.factor = TRUE
              )

names(df_study1_long) #check names of long dataset
dim(df_study1_long) #check dimensions of the dataframe
# Original dataset with 6 waves
df_study2_long <- df_study1_long

@@

df_study1_long <- 
set_label(x = df_study1_long,
          label = get_label(select(df_study1,names(df_study1_long))))

@@

names(df_study1_long_t7)
dim(df_study1_long_t7)
sjmisc::frq(df_study1_long_t7$ola)

# SAVE DATA____________________________________________________________________
save(df_study1_long,file = here::here("input/data-proc/df_study1_long.RData"))
save(df_study1_long_t7,file = here::here("input/data-proc/df_study1_long_t7.RData"))
save(df_study2_long,file = here::here("input/data-proc/df_study2_long.RData"))
# save codebook
sjPlot::view_df(df_study1_long,
        show.frq = T,
        show.prc = T,
        show.na = T,
        file = here::here("output/df_study1_long.html"))

>>>>

```{r}
remove(list = ls()) # clean workspace
# .rs.restartR() #restart R session
if (!require("pacman")) install.packages("pacman") # install pacman
# load libraries
pacman::p_load(dplyr,readxl,sjmisc, sjlabelled, questionr, car,here, ggstats)
load(here::here("input/data-original/elsoc_long.Rdata"))

@@

# Attitudes toward inmigrants__________________________________________________
elsoc_long_2016_2023 <- 
elsoc_long_2016_2023 %>% 
  #create new variables
  mutate(just_educ=d02_02,
         just_pension=d02_01,
         just_salud=d02_03,
         des_perc=c18_11,
         sexo=m0_sexo,
         merit_effort=c18_09,
         merit_talent=c18_10,
         ahead_family=r12_02,
         ahead_educ=r12_03,
         ahead_ambition=r12_04,
         ahead_hardwork=r12_05,
         covid_conocidos=r03_09,
         covid_economia=c37_09,
         covid_cuarentena=c37_10,
         covid_propio=s31,
         covid_alimentos=m62,
         covid_beneficios=m63_01,
         covid_retirouno=m63_02,
         covid_retirodos=m63_03
         ) %>% 
  sjlabelled::drop_labels(., drop.na = FALSE)
```

```{r}
#
sjmisc::frq(elsoc_long_2016_2023$just_educ)
elsoc_long_2016_2023$just_educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$just_educ, 
                      label = "Education distributive justice")

@@

sjmisc::frq(elsoc_long_2016_2023$just_educ)
sjmisc::frq(elsoc_long_2016_2023$merit_effort)
sjmisc::frq(elsoc_long_2016_2023$merit_talent)
sjmisc::frq(elsoc_long_2016_2023$ahead_family)
sjmisc::frq(elsoc_long_2016_2023$ahead_educ)
sjmisc::frq(elsoc_long_2016_2023$ahead_ambition)
sjmisc::frq(elsoc_long_2016_2023$ahead_hardwork)
```

```{r}
#Age Ranges
elsoc_long_2016_2023$edad <- 
  factor(car::recode(elsoc_long_2016_2023$m0_edad, 
                     "18:29=1;30:49=2;50:64=3;65:150=4"),
         labels = c('18-29', '30-49', '50-64', '65 or more'))
elsoc_long_2016_2023$edad <-
  sjlabelled::set_label(elsoc_long_2016_2023$edad, 
                        label = c("Age groups")) 
# Other controls______________________________________________________________
# Subjective social status: individual ____________________________

@@

```{r}
# Socioeconomic_________________________________________________________________

# Education_______________________________________
elsoc_long_2016_2023$educ <- 
  car::recode(elsoc_long_2016_2023$m01,
              "c(1,2,3,4,5,6,7)=1;c(8,9,10)=2; c(-888,-999)=NA")
elsoc_long_2016_2023$educ <-
  factor(elsoc_long_2016_2023$educ,
         labels = c("Less than Universitary","Universitary"))

class(elsoc_long_2016_2023$educyear)
sjmisc::frq(elsoc_long_2016_2023$educyear)
```

```{r}
# N Household:
# Select variables______________________________________________________________

df_study1 <- 
elsoc_long_2016_2023 %>% 
  filter(muestra == 1) %>% #keep original sample (vs refresh)
  select(idencuesta,ola,region,region_cod,comuna,comunacod=comuna_cod,
         muestra,
         ponderador_long_total, segmento, estrato,
         sexo,
         just_educ,
         just_pension,
         just_salud,
         des_perc,
         merit_effort,
         merit_talent,
         ahead_family,
         ahead_educ,
         ahead_ambition,
         ahead_hardwork,
         educ,educyear,
         # ing_pc,
         # quintil,
#         quintil1,
         ess,
         covid_conocidos,
         covid_economia,
         covid_cuarentena,
         covid_propio,
         covid_alimentos,
         covid_beneficios,
         covid_retirouno,
         covid_retirodos,
         pos_id,
         edad,
         nhogar1,
         m46_nhogar,
         m54, m30, m30b, m29)

#Impute midpoint of income ranges
df_study1$m30_rec <-
as.numeric(car::recode(df_study1$m30,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

#Impute midpoint of income ranges (2021)
df_study1$m30b_rec <-
as.numeric(car::recode(df_study1$m30b,
           "1=125000;2=300000;3=400000;4=575000;5=70000;NA=NA;c(-888,-999)=NA"))

sjmisc::frq(df_study1$m30b_rec)

#Recode DK/DA of Income to NA
df_study1$m29_rec <-
  as.numeric(car::recode(df_study1$m29,"c(-888,-999)=NA"))

#replace NA of income with new imputed variable
df_study1$m29_imp <- 
  ifelse(test = !is.na(df_study1$m29_rec),
         yes =  df_study1$m29_rec,
         no =  df_study1$m30_rec)
summary(df_study1$m29_imp)

df_study1$m29_imp <- 
  ifelse(test = is.na(df_study1$m29_imp),
         yes =  df_study1$m30b_rec,
         no =  df_study1$m29_imp)
summary(df_study1$m29_imp)

# deflate at each year's prices
df_study1$deflactor <-
  with(df_study1, case_when(
    ola == 2016 ~ 113.88 / 123.82,
    ola == 2017 ~ 116.46 / 123.82,
    ola == 2018 ~ 119.45 / 123.82,
    ola == 2019 ~ 123.82 / 123.82
  ))

# Reshape long to wide
df_study1_wide <- df_study1 %>% 
  tidyr::pivot_wider(id_cols = c("idencuesta","muestra"),
                     names_from = "ola",
                     # names_prefix = c("educ","educyear","ing_pc","quintil","quintil1"),
                     values_from = names(select(df_study1,comuna,comunacod,ponderador_long_total,segmento, estrato, just_educ:edad,sexo, nhogar1, m46_nhogar, m54, m30, m30b, m29_imp))
                     )

df_study1_wide$m54_6 <- df_study1_wide$m54_7

# reshape from long to wide
pacman::p_load(datasets,data.table)
df_study1_long <- data.table::melt.data.table(data.table::setDT(df_study1_wide),
              id.vars = c("idencuesta","muestra"),
              variable.name = c("ola"),
              measure = patterns("^comuna_","^comunacod_","^ponderador_long_total", "^segmento", "^estrato",
                                 "^just_educ_", "^just_pension_", "^just_salud_", "^des_perc_",
                                 "^merit_effort_","^merit_talent_",
                                 "^ahead_family_", "^ahead_educ_", "^ahead_ambition_", "^ahead_hardwork_",
                                 "^covid_conocidos_", "^covid_economia_", "^covid_cuarentena_", "^covid_propio_",
                                 "^covid_alimentos_", "^covid_beneficios_", "^covid_retirouno_", "^covid_retirodos_",
                                 "^educ_","^educyear_","^ess_",
                                 "^pos_id_",
                                 "^edad_","^sexo_", "^m54_", "^nhogar1_", "^m46_nhogar_", "^m30_", "^m30b_", "^m29_imp_"
                                 ),
              value.name = c("comuna","comunacod","ponderador_long_total","segmento","estrato",
                             "just_educ", "just_pension","just_salud", "des_perc",
                             "merit_effort","merit_talent",
                             "ahead_family","ahead_educ","ahead_ambition","ahead_hardwork",
                             "covid_conocidos", "covid_economia", "covid_cuarentena", "covid_propio",
                             "covid_alimentos", "covid_beneficios", "covid_retirouno", "covid_retirodos",
                             "educ","educyear","ess",
                             "pos_id",
                             "edad","sexo", "m54", "nhogar1", "m46_nhogar", "m30", "m30b", "m29_imp"),
              na.rm = F,value.factor = TRUE
              )

df_study1_long <-
  df_study1_long %>%
  mutate(n_hogar =
           dplyr::case_when(ola == 1 ~ nhogar1,
                            ola == 2 ~ m46_nhogar,
                            ola == 3 ~ m54,
                            ola == 4 ~ m54,
                            ola == 5 ~ m54,
                            ola == 6 ~ m54,
                            ola == 7 ~ m54))
sjmisc::frq(df_study1_long$n_hogar)

#Recode DK/DA to NA
df_study1_long$n_hogar_r<-
  car::recode(df_study1_long$n_hogar,"c(-888,-999)=NA")

# Per capita household income:
df_study1_long$ing_pc <- 
  (df_study1_long$m29_imp/df_study1_long$n_hogar_r)

df_study1_long$ing_pc <-
sjlabelled::set_label(x = df_study1_long$ing_pc,
                      label = "Household income per capita")

sjmisc::descr(df_study1_long$ing_pc)

# Compute income quintiles
df_study1_long <- df_study1_long %>% 
  group_by(ola) %>% 
  mutate(quintil = ntile(-desc(ing_pc), 5)) %>% 
  ungroup()

df_study1_long$quintil <- 
  factor(df_study1_long$quintil,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

df_study1_long$quintil <- 
sjlabelled::set_label(x = df_study1_long$quintil,
                      label = "Household income quintile per capita")

sjmisc::frq(df_study1_long$quintil)

#include new quintile category with missing cases
df_study1_long$quintil1<-
  car::recode(df_study1_long$quintil, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

df_study1_long$quintil1 <- 
sjlabelled::set_label(x = df_study1_long$quintil1,
                      label = "Household income quintile per capita (NA)") 
sjmisc::frq(df_study1_long$quintil1)
```

# Alternative operationalization
_Income_
```{r}
df_study1_long <-
df_study1_long %>%
  # group_by(ola) %>%
  mutate(percentile_wave=ntile(ing_pc,100),
         bot_mid_top_wave=car::recode(percentile_wave,"1:20=1;21:79=2;80:100=3"))

elsoc_wide <- panelr::widen_panel(panelr::panel_data(df_study1_long %>% select(idencuesta,ola,percentile_wave,bot_mid_top_wave), id = idencuesta, wave = ola), separator = "_")
elsoc_wide$income_prof <- paste0(elsoc_wide$bot_mid_top_wave_1,elsoc_wide$bot_mid_top_wave_5)
frq(elsoc_wide$income_prof)
# a. Stable low income: In the lowest tertile in all three periods
# b. Stable middle income: In the middle tertile in all three periods
# c. Stable high income: In the highest tertile in all three periods
# d. Upward mobility: In the lowest tertile in the first period and either in tertile 2 or 3 in the last period
# e. Downward mobility: In the highest tertile in the first period and either in tertile 1 or 2 in the last period
# f. Other: All other trajectories
elsoc_wide$ind_inc_prof <- car::recode(elsoc_wide$income_prof,"'11'='1stablelow';'33'='2stablemid';'22'='3stablehig';c('12','13')='4upwardmob';c('31','21')='5downwardmob';else = '99other'")
frq(elsoc_wide$ind_inc_prof)
# • Stable low income: Start and end at the 20th percentile in income (P20).
# • Stable middle income: Start and end at the 50th percentile in income (P50).
# • Stable high income: Start and end at the 80th percentile in income (P80).
# • Upward mobility: Start at 20th percentile, end at 80th percentile (P20 → P80).
# • Downward mobility: Start at 80th percentile, end at 20th percentile (P80 → P20).
df_profiles<- select(elsoc_wide,"idencuesta","ind_inc_prof",percentile_wave_1,percentile_wave_5)
df_profiles$idencuesta <- as.numeric(as.character(df_profiles$idencuesta))
df_study1_long <-
  df_study1_long %>%
  left_join(select(df_profiles,"idencuesta","ind_inc_prof",percentile_wave_1,percentile_wave_5),by="idencuesta")
```

```{r}
# Income group as percentage of median
# poverty (<50 per cent of median income), below average (>¼50 per cent and <100 per cent), above average (>¼100 per cent and <200 per cent), and affluence (>¼200 per cent).
df_study1_long<-
  df_study1_long %>%
  group_by(ola) %>%
  #compute median household pc income
  mutate(inc_pcmed=weighted.median(ing_pc,ponderador_long_total,na.rm = T)) %>%
  mutate(ratioincome=(ing_pc/inc_pcmed)*100) %>%
  ungroup()
# elsoc_long_2016_2022.2[,c('idencuesta',"ola", 'm29_imp',"n_hogar","ing_pc","inc_pcmed")] %>% View()
df_study1_long$incomegrp=
  car::recode(df_study1_long$ratioincome,
              "0:49.9='Low';50:99.9='Below average';100:200='Above average';200.1:999999999999='Affluence'",
              levels =c("Low","Below average","Above average","Affluence"),as.factor = T)
frq(df_study1_long$incomegrp)
```

```{r}
# Reshape long to wide
df_study1_wide <- df_study1_long %>% 
  tidyr::pivot_wider(id_cols = c("idencuesta","muestra"),
                     names_from = "ola",
                     # names_prefix = c("educ","educyear","ing_pc","quintil","quintil1"),
                     values_from = names(select(df_study1_long,comuna,comunacod,ponderador_long_total,segmento, estrato, just_educ:edad,sexo, quintil1, percentile_wave, bot_mid_top_wave, inc_pcmed, ratioincome, incomegrp))
                     )

@@

sjPlot::view_df(df_study1_wide,
                show.frq = T,show.values = T,show.na = T,show.prc = T, show.type = T)
```

```{r}
# reshape from long to wide
pacman::p_load(datasets,data.table)
df_study1_long <- data.table::melt.data.table(data.table::setDT(df_study1_wide),
              id.vars = c("idencuesta","muestra"),
              variable.name = c("ola"),
              measure = patterns("^comuna_","^comunacod_","^ponderador_long_total", "^segmento", "^estrato",
                                 "^just_educ_", "^just_pension_", "^just_salud_", "^des_perc_",
                                 "^merit_effort_","^merit_talent_",
                                 "^ahead_family_", "^ahead_educ_", "^ahead_ambition_", "^ahead_hardwork_",
                                 "^covid_conocidos_", "^covid_economia_", "^covid_cuarentena_", "^covid_propio_",
                                 "^covid_alimentos_", "^covid_beneficios_", "^covid_retirouno_", "^covid_retirodos_",
                                 "^educ_","^educyear_","^quintil1_","^ess_",
                                 "^pos_id_",
                                 "^edad_","^sexo_",
                                 "^percentile_wave_", "^bot_mid_top_wave_", "^inc_pcmed_", "^ratioincome_", "^incomegrp_"
                                 ),
              value.name = c("comuna","comunacod","ponderador_long_total","segmento","estrato",
                             "just_educ", "just_pension","just_salud", "des_perc",
                             "merit_effort","merit_talent",
                             "ahead_family","ahead_educ","ahead_ambition","ahead_hardwork",
                             "covid_conocidos", "covid_economia", "covid_cuarentena", "covid_propio",
                             "covid_alimentos", "covid_beneficios", "covid_retirouno", "covid_retirodos",
                             "educ","educyear","quintil1","ess",
                             "pos_id",
                             "edad","sexo",
                             "percentile_wave", "bot_mid_top_wave", "inc_pcmed", "ratioincome", "incomegrp"),
              na.rm = F,value.factor = TRUE
              )
```

```{r}
names(df_study1_long) #check names of long dataset
dim(df_study1_long) #check dimensions of the dataframe
# Original dataset with 6 waves
df_study2_long <- df_study1_long

@@

# df_study1_long <- 
# set_label(x = df_study1_long,
#           label = get_label(select(df_study1,names(df_study1_long))))

@@

names(df_study1_long_t7)
dim(df_study1_long_t7)
sjmisc::frq(df_study1_long_t7$ola)
```

```{r}
# SAVE DATA____________________________________________________________________
save(df_study1_long,file = here::here("input/data-proc/df_study1_long.RData"))
save(df_study1_long_t7,file = here::here("input/data-proc/df_study1_long_t7.RData"))
save(df_study2_long,file = here::here("input/data-proc/df_study2_long.RData"))
# save codebook
sjPlot::view_df(df_study1_long,
        show.frq = T,
        show.prc = T,
        show.na = T,
        file = here::here("output/df_study1_long.html"))
